---
title: "Assignment-B1"
author: "Tongshuai Zhang"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1: Make a Function

In the MDA project, I repeated use the used **group_by() %>% summarise()** statement for **vancouver_trees** dataset Therefore, it would be appropriate 
to create a generic function that group by a specific column and summarize the proportion calculated by another specific column in the dataset.

```{r}
library(dplyr)
library(datateachr) # <- might contain the data you picked!
library(tidyverse)

group_distinct_col_count_prop <- function(dataset, group_col, summary_col) {
  stopifnot(is.data.frame(dataset))
  
  # Check if group_col and summary_col are present in the dataset
  if (!(group_col %in% colnames(dataset))) {
    stop("Error: group_col is not a column in the dataset.")
  }
  if (!(summary_col %in% colnames(dataset))) {
    stop("Error: summary_col is not a column in the dataset.")
  }
  
  total_summary_col_distinct_count <- n_distinct(dataset[[summary_col]])
  
  result <- dataset %>%
    group_by(.data[[group_col]]) %>%
    summarise(total = n_distinct(.data[[summary_col]])) %>%
    mutate(proportion = total / total_summary_col_distinct_count)
  return(result)
}
```

# Exercise 2: Document the Function

```{r}

#' @title Group by a Column and Calculate Proportion of Distinct Values in Another Column
#' @description 
#' This function groups a given dataset by a specified column and then calculates the  
#' proportion of distinct values in another specified column within each group. 
#' It is designed to provide a quick and generic way to summarize data, which is useful in exploratory data analysis.
#' @param dataset: A data frame containing the data to be grouped and summarized.
#' @param group_col: A string representing the name of the column to group by.
#' @param summary_col: A string representing the name of the column to calculate the proportion of distinct values.
#' @return: A data frame containing the grouping column, the total count of distinct values in the summary_col for each group, and 
#' the proportion of these distinct values relative to the total distinct values in summary_col across the entire dataset.

group_distinct_col_count_prop <- function(dataset, group_col, summary_col) {
  stopifnot(is.data.frame(dataset))
  
  # Check if group_col and summary_col are present in the dataset
  if (!(group_col %in% colnames(dataset))) {
    stop("Error: group_col is not a column in the dataset.")
  }
  if (!(summary_col %in% colnames(dataset))) {
    stop("Error: summary_col is not a column in the dataset.")
  }
  
  total_summary_col_distinct_count <- n_distinct(dataset[[summary_col]], na.rm = TRUE)
  
  result <- dataset %>%
    group_by(.data[[group_col]]) %>%
    summarise(total = n_distinct(.data[[summary_col]], na.rm = TRUE)) %>%
    mutate(proportion = total / total_summary_col_distinct_count)
  return(result)
}
```

# Exercise 3: Include examples

## Example 01:
```{r}
# Check the number of species and its proportion to the total species for every distinct neighbourhood_name
vancouver_trees_summary_by_neighbourhood <- group_distinct_col_count_prop(vancouver_trees, group_col = "neighbourhood_name", summary_col = "species_name")
vancouver_trees_summary_by_neighbourhood
```

## Example 02:
```{r}
# Check the number of species and its proportion to the total species for every distinct std_street
vancouver_trees_summary_by_str_street <- group_distinct_col_count_prop(vancouver_trees, group_col = "std_street", summary_col = "species_name")
vancouver_trees_summary_by_str_street
```

## Example 03:
```{r}
vancouver_trees
# Check the number of species and its proportion to the total species for every distinct genus
vancouver_trees_summary_by_genus <- group_distinct_col_count_prop(vancouver_trees, group_col = "genus_name", summary_col = "species_name")
vancouver_trees_summary_by_genus
```

# Exercise 4: Test the Function

```{r}
library(testthat)
# Define the function group_distinct_col_count_prop here

# Sample test data
test_data <- data.frame(
  Species = c("Oak", "Maple", "Oak", "Pine", "Maple", "Pine", NA),
  Height = c(10, 15, 20, 25, 30, 35, 40)
)

# Test cases
test_that("group_distinct_col_count_prop function works correctly", {
  
  # Test 1: Vector with no NA's
  result1 <- group_distinct_col_count_prop(test_data[1:6, ], "Species", "Height")
  result1
  expect_equal(nrow(result1), 3)
  expect_equal(sum(result1$proportion), 1)
  
  # Test 2: Vector that has NA's
  result2 <- group_distinct_col_count_prop(test_data, "Species", "Height")
  expect_equal(nrow(result2), 4)
  expect_equal(sum(result2$proportion, na.rm = TRUE), 1)
  
  # Test 3: Vector of length 0
  empty_data <- test_data[FALSE, ]
  result3 <- group_distinct_col_count_prop(empty_data, "Species", "Height")
  expect_equal(nrow(result3), 0)
})

# Test 4: Vector that hasn't the column specified in the function arguments
test_that("group_distinct_col_count_prop handles non-existent columns correctly", {
  expect_error(group_distinct_col_count_prop(test_data, "Species", "Heights"),
               "Error: summary_col is not a column in the dataset.")
})
```